Cell Decomposition of Rank 2 Quiver Grassmannians

### 2017-11-16 Grassmannian_cells.tex
insert result of Reineke et al

\newcommand{\bfe}{\mathbf{e}}
\newcommand{\bff}{\mathbf{f}}
\newcommand{\bfg}{\mathbf{g}}
\newcommand{\bfi}{\mathbf{i}}
\newcommand{\bfh}{\mathbf{h}}
\newcommand{\tbfe}{{\tilde\bfe}}
\newcommand{\tbff}{{\tilde\bff}}
\newcommand{\tbfg}{{\tilde\bfg}}
\newcommand{\tbfh}{{\tilde\bfh}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\cC}{\mathcal{C}}
\newcommand{\cP}{\mathcal{P}}
\newcommand{\ui}{\underline i}
\newcommand{\uj}{\underline j}
\newcommand\udim{{\underline{\dim}\, }}
\newcommand{\pt}{\mathrm{pt}}
\newcommand{\Rep}{\mathrm{Rep}}
\newcommand{\rep}{\operatorname{rep}}
\newcommand{\Gr}{\mathrm{Gr}}
\renewcommand{\AA}{\mathbb{A}}
\newcommand{\CC}{\mathbb{C}}
\newcommand{\kk}{\Bbbk}
\newcommand{\ZZ}{\mathbb{Z}}
\newcommand{\Ext}{\operatorname{Ext}}
\newcommand{\Hom}{\operatorname{Hom}}
\newcommand{\im}{\operatorname{im}}
\newcommand{\codim}{\operatorname{codim}}
\newcommand{\into}{\hookrightarrow}
\newcommand{\onto}{\to\!\!\!\!\!\to}
\newcommand{\ses}[3]{0\rightarrow #1\rightarrow #2\rightarrow#3\rightarrow 0}
\newcommand{\Sc}[2]{\langle #1,#2\rangle}
\newcommand{\spn}{\operatorname{span}}
\setcounter{MaxMatrixCols}{20}
\title{Cell Decomposition of Rank 2 Quiver Grassmannians}

@@

In particular, we have $\chi(X)=\chi(X^{\C^\ast})$ for the Euler characteristic.
\end{theorem}

\section{Torus Actions on Quiver Grassmannians}\label{torusaction}
\noindent 
Fix a vector space $X$ of dimension $n$ and let $d\leq n$. We first consider a natural torus action on usual Grassmannians coming along with any direct sum decomposition of the vector space $X$. Then we generalize the concept to quiver Grassmannians, but we will see that the torus fixed point set can be calculated in an analogous manner. Thus assume that $X=\bigoplus_{j=1}^l X_j$ is a direct sum decomposition of $X$. We choose a basis $\mathcal B=\{v_1,\ldots,v_n\}$ of $X$ which is compatible with this decomposition, i.e. there exist numbers $1\leq j_1\leq j_2\ldots\leq j_{l-1}\leq n$ such that
\[v_1,\ldots,v_{j_1}\in X_1,\,v_{j_1+1},\ldots,v_{j_2}\in X_2,\ldots,v_{j_{l-1}+1},\ldots,v_{n}\in X_l\]
Moreover, we choose a map $d:\{1,\ldots,n\}\to\mathbb{Z}$ such that $d(i)=d(i')$ if $v_i,v_{i'}\in X_{j}$ for some $j$. Then there exists a $\C^\ast$-action on $X$ when extending the definition $t.v_i:=t^{d(j)}v_i$ if $v_i\in X_j$ linearly to $X$. This naturally induces an action of $\C^\ast$ on the usual Grassmannians $\Gr_d(X)$. We can write every $U\in\Gr_d(X)$ uniquely in row-echelon form with respect to the basis $\mathcal B$, i.e. there exists a unique sequence $1\leq i_1\leq\ldots\leq i_d\leq n$ and a unique matrix representation of $U$ of the form
\[M(U):=\begin{pmatrix}\ast&\ldots &\ast &1&0 &\ldots&\ldots&\ldots&\ldots&\ldots&\ldots&\ldots&\ldots&\ldots&0\\\ast&\ldots &\ast&0&\ast&\ldots &\ast&1&0&\ldots&\ldots&\ldots&\ldots&\ldots&0\\\vdots &\ddots&\vdots&\vdots&\vdots &\ddots&\vdots&0&\ast&\ddots&\ddots&\ddots&\ddots&\ddots&\vdots\\\vdots &\ddots&\vdots&\vdots&\vdots &\ddots&\vdots&\vdots&\ddots&\ddots&\ddots&\ddots&\ddots&\ddots&\vdots\\\ast&\ldots &\ast&0&\ast&\ldots &\ast&0&\ast&\ldots&\ast&1&0&\ldots&0\end{pmatrix}\in M_{d,n}(k)\]
where the unit vectors are in the columns $i_1,\ldots,i_d$. All matrix representations of this form define a Schubert cell $X_{\bfi}$ where $\bfi=(i_1,\ldots,i_d)$. The $\C^\ast$-action is then given by multiplying the $w^{\mathrm{th}}$ column with $t^{d(w)}$, i.e. we have
\[(t.M(U))_{u,w}=t^{d(w)}\cdot M(U)_{u,w}.\]
Every Schubert cell is invariant under the torus action, i.e. $t.M(U)\in X_\bfi$ for all $M(U)\in X_{\bfi}$. 
\begin{lemma}\label{usualGrass}
Choose $d$ in such a way that $d(i)\neq d(i')$ if $v_i\in X_j$, $v_{i'}\in X_{j'}$ and $j\neq j'$.
Then we have $U\in X_{\bfi}^{\C^\ast}$ if and only if $U=\bigoplus_{i=1}^l U\cap X_i$.
\end{lemma}
\begin{proof}If $U=\bigoplus_{i=1}^l U\cap X_i$, we have 
\[t.U=\bigoplus_{i=1}^l t.(U\cap X_i)=\bigoplus_{i=1}^l t^{d(j_i+1)}(U\cap X_i)=U\]
where $j_0=0$.

@@

By Lemma \ref{directsums} and induction hypothesis every quiver Grassmannian on the right hand side has a cell decomposition. The second part of the Proposition says that the fibres are empty if and only if $\tbfg=0$ and $\tbfh=\udim \tilde P_{m+1}^{(k)}$. In all other cases, the fiber dimension only depend on the dimension vectors $\tbff, \tbfg,\tbfh$ which means that they are constant over $Gr^{\tilde Q}_\tbff( P(m,k))\times Gr^{\tilde Q}_\tbfg(\tau\tilde P_{m+1}^{(k)})\times Gr^{\tilde Q}_\tbfh(\tilde P_{m+1}^{(k)})$\footnote{actually this should already say that this is a vector bundle as the base space is smooth (all reps are exceptional)}. Analogously to Lemma \ref{directsums}, we can conclude that this induces trivial bundles over every affine cell. In total, this yields a cell decomposition of the quiver Grassmannian $Gr^{\tilde Q}_\tbfe(\tilde P_{m+1}^{(k-1)})$.

@@

\begin{thebibliography}{10}
\bibitem{ass}
Assem, I., Simson, D., Skowronski, A.: Elements of the Representation Theory of Associative Algebras. Cambridge University Press, Cambridge 2007.
\bibitem{bb} Bialynicki-Birula, A.: Some theorems on actions of algebraic groups. Annals of Mathematics \textbf{98}, 480-497 (1973).
\bibitem{bgp}
Bernstein, I.~N., Gelfand, I.~M., Ponomarev, V.~A.: Coxeter functors, and Gabriel's theorem. Russian Mathematical Surveys \textbf{28}(2), 17-32 (1973).
\bibitem{brenner-butler} S. Benner, M.~C.~R. Butler, The equivalence of certain functors occuring in the representation theory of artin algebras and species, J. London Math. Soc., 14 (1976), 183-187.
\bibitem{cc}
Caldero, P., Chapoton, F.: Cluster algebras as {H}all algebras of quiver representations.
Commentarii Mathematici Helvetici \textbf{81}(3), 595-616 (2006).
\bibitem{ck}
  P. Caldero, B. Keller: From triangulated categories to cluster algebras II.  Ann. Sci. \'Ecole Norm. Sup. (4) \textbf{39} (2006), no. 6, pp.~983--1009.
\bibitem{cr}
Caldero, P., Reineke, M.: On the quiver Grassmannian in the acyclic case.
Journal of Pure and Applied Algebra \textbf{212}(11), 2369-2380 (2008).
\bibitem{cj} Crawley-Boevey, W., Jensen, Bernt Tore: A note on sub-bundles of vector bundles. Glasgow Mathematical Journal \textbf{48}, 459-462 (2006).

\bibitem{gab} Gabriel, P.: The universal cover of a finite-dimensional algebra. Representations of algebras. Lecture Notes in Mathematics {\bf 903}, 68-105 (1981).
\bibitem{hr} Happel, D., Ringel, C.M.: Tilted Algebras. Transactions of the American Mathematical Society {\bf 274}, no.2, 399-443 (1982).
\bibitem{llz} K. Lee, L. Li, A. Zelevinsky: Greedy elements in rank 2 cluster algebras. Selecta Math. \textbf{20} (2014) pp.~57--82.
\bibitem{rin} Ringel, C.M.: Reflection functors for hereditary algebras. Journal of the London Mathematical Society (2) {\bf 21}, no. 3, 465-479 (1980).

\bibitem{rupel} Rupel, D.: Rank Two Non-Commutative Laurent Phenomenon and Pseudo-Positivity, arXiv.

>>>>

\newcommand{\bfe}{\mathbf{e}}
\newcommand{\bff}{\mathbf{f}}
\newcommand{\bfg}{\mathbf{g}}
\newcommand{\bfi}{\mathbf{i}}
\newcommand{\bfh}{\mathbf{h}}
\newcommand{\tbfe}{{\tilde\bfe}}
\newcommand{\tbff}{{\tilde\bff}}
\newcommand{\tbfg}{{\tilde\bfg}}
\newcommand{\tbfh}{{\tilde\bfh}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\cC}{\mathcal{C}}
\newcommand{\cG}{\mathcal{G}}
\newcommand{\cP}{\mathcal{P}}
\newcommand{\ui}{\underline i}
\newcommand{\uj}{\underline j}
\newcommand\udim{{\underline{\dim}\, }}
\newcommand{\pt}{\mathrm{pt}}
\newcommand{\Rep}{\mathrm{Rep}}
\newcommand{\rep}{\operatorname{rep}}
\newcommand{\Gr}{\mathrm{Gr}}
\renewcommand{\AA}{\mathbb{A}}
\newcommand{\CC}{\mathbb{C}}
\newcommand{\kk}{\Bbbk}
\newcommand{\ZZ}{\mathbb{Z}}
\newcommand{\Ext}{\operatorname{Ext}}
\newcommand{\Hom}{\operatorname{Hom}}
\newcommand{\im}{\operatorname{im}}
\newcommand{\codim}{\operatorname{codim}}
\newcommand{\into}{\hookrightarrow}
\newcommand{\onto}{\to\!\!\!\!\!\to}
\newcommand{\ses}[3]{0\rightarrow #1\rightarrow #2\rightarrow#3\rightarrow 0}
\newcommand{\Sc}[2]{\langle #1,#2\rangle}
\newcommand{\spn}{\operatorname{span}}
\setcounter{MaxMatrixCols}{20}
\title{Cell Decomposition of Rank 2 Quiver Grassmannians}

@@

In particular, we have $\chi(X)=\chi(X^{\C^\ast})$ for the Euler characteristic.
\end{theorem}
Let $\ses{M}{B}{N}$ be a short exact sequence of representations. This induces a map between quiver Grassmannians
\[\Psi_\tbfe:Gr_\tbfe(B)\to\bigsqcup_{\tbff+\tbfg=\tbfe}  Gr_\tbff(M)\times Gr_\tbfg(N),\,U\mapsto (U\cap M,(U+M)/M)\]
which is also called Caldero-Chapoton map.  By the results of \cite[Section 3]{cc} it follows that $\Psi^{-1}_\tbfe(A,V)=\mathbb{A}^{\dim\Hom(A,M/V)}$ for all pairs of subrepresentations $(A,V)$ for which the fibre is not empty.
Let $\cG_{\tbff,\tbfg}:=\Psi^{-1}_\tbfe(\Gr_\tbff(M)\times \Gr_\tbfg(N))$. Then we have
$$\Gr_\tbfe(B)=\bigsqcup_{\tbff+\tbfg=\tbfe} \cG_{\tbff,\tbfg}.$$
The map $\Psi_\tbfe$ restricts to morphisms of projective varieties
\[\Psi_\tbfe^{\tbff,\tbfg}:\cG_{\tbff,\tbfg}\to \Gr_\tbff(M)\times \Gr_\tbfg(N).\]

The following is proved in \cite[Section 3]{cefr}.
\begin{theorem}\label{vb}
If the fibre dimension of $\Psi_\tbfe^{\tbff,\tbfg}$ is constant over $\mathrm{Im}(\Psi_\tbfe^{\tbff,\tbfg})$, then it an affine bundle over the image. In particular, the existence of a cell decomposition of $\mathrm{Im}(\Psi_\tbfe^{\tbff,\tbfg})$ implies a cell decomposition of $\cG_{\tbff,\tbfg}$ in this case.
\end{theorem}
\begin{proof}\footnote{sketch, double-check}

As $X$ and $Y$ are projective, they are compact. As the fibres are affine spaces of constant dimension and as $X$ and $Y$ are smooth, the induced morphism on the tangent spaces is surjective for all points $x\in X$. This shows that $f$ is smooth of relative dimension $n$, see Hartshorne and thus a submersion?!

\end{proof}
\section{Torus Actions on Quiver Grassmannians}\label{torusaction}
\noindent 
Fix a vector space $X$ of dimension $n$ and let $d\leq n$. We first consider a natural torus action on usual Grassmannians coming along with any direct sum decomposition of the vector space $X$. Then we generalize the concept to quiver Grassmannians, but we will see that the torus fixed point set can be calculated in an analogous manner. Thus assume that $X=\bigoplus_{j=1}^l X_j$ is a direct sum decomposition of $X$. We choose a basis $\mathcal B=\{v_1,\ldots,v_n\}$ of $X$ which is compatible with this decomposition, i.e. there exist numbers $1\leq j_1\leq j_2\ldots\leq j_{l-1}\leq n$ such that
\[v_1,\ldots,v_{j_1}\in X_1,\,v_{j_1+1},\ldots,v_{j_2}\in X_2,\ldots,v_{j_{l-1}+1},\ldots,v_{n}\in X_l\]
Moreover, we choose a map $d:\{1,\ldots,n\}\to\mathbb{Z}$ such that $d(i)=d(i')$ if $v_i,v_{i'}\in X_{j}$ for some $j$. Then there exists a $\C^\ast$-action on $X$ when extending the definition $t.v_i:=t^{d(j)}v_i$ if $v_i\in X_j$ linearly to $X$. This naturally induces an action of $\C^\ast$ on the usual Grassmannians $\Gr_d(X)$. We can write every $U\in\Gr_d(X)$ uniquely in row-echelon form with respect to the basis $\mathcal B$, i.e. there exists a unique sequence $1\leq i_1\leq\ldots\leq i_d\leq n$ and a unique matrix representation of $U$ of the form
\[M(U):=\begin{pmatrix}\ast&\ldots &\ast &1&0 &\ldots&\ldots&\ldots&\ldots&\ldots&\ldots&\ldots&\ldots&\ldots&0\\\ast&\ldots &\ast&0&\ast&\ldots &\ast&1&0&\ldots&\ldots&\ldots&\ldots&\ldots&0\\\vdots &\ddots&\vdots&\vdots&\vdots &\ddots&\vdots&0&\ast&\ddots&\ddots&\ddots&\ddots&\ddots&\vdots\\\vdots &\ddots&\vdots&\vdots&\vdots &\ddots&\vdots&\vdots&\ddots&\ddots&\ddots&\ddots&\ddots&\ddots&\vdots\\\ast&\ldots &\ast&0&\ast&\ldots &\ast&0&\ast&\ldots&\ast&1&0&\ldots&0\end{pmatrix}\in M_{d,n}(k)\]
where the unit vectors are in the columns $i_1,\ldots,i_d$. All matrix representations of this form define a Schubert cell $X_{\bfi}$ where $\bfi=(i_1,\ldots,i_d)$. The $\C^\ast$-action is then given by multiplying the $w^{\mathrm{th}}$ column with $t^{d(w)}$, i.e. we have
\[(t.M(U))_{u,w}=t^{d(w)}\cdot M(U)_{u,w}.\]
Every Schubert cell is invariant under the torus action, i.e. $t.M(U)\in X_\bfi$ for all $M(U)\in X_{\bfi}$. 
\begin{lemma}\label{usualGrass}
Choose $d$ in such a way that $d(i)\neq d(i')$ if $v_i\in X_j$, $v_{i'}\in X_{j'}$ and $j\neq j'$.
Then we have $U\in X_{\bfi}^{\C^\ast}$ if and only if $U=\bigoplus_{i=1}^l U\cap X_i$.
\end{lemma}
\begin{proof}If $U=\bigoplus_{i=1}^l U\cap X_i$, we have 
\[t.U=\bigoplus_{i=1}^l t.(U\cap X_i)=\bigoplus_{i=1}^l t^{d(j_i+1)}(U\cap X_i)=U\]
where $j_0=0$.

@@

By Lemma \ref{directsums} and induction hypothesis every quiver Grassmannian on the right hand side has a cell decomposition. The second part of the Proposition says that the fibres are empty if and only if $\tbfg=0$ and $\tbfh=\udim \tilde P_{m+1}^{(k)}$. In all other cases, the fiber dimension only depend on the dimension vectors $\tbff, \tbfg,\tbfh$ which means that they are constant over $Gr^{\tilde Q}_\tbff( P(m,k))\times Gr^{\tilde Q}_\tbfg(\tau\tilde P_{m+1}^{(k)})\times Gr^{\tilde Q}_\tbfh(\tilde P_{m+1}^{(k)})$\footnote{actually this should already say that this is a vector bundle as the base space is smooth (all reps are exceptional), apply Lemma \ref{vb}?}. Analogously to Lemma \ref{directsums}, we can conclude that this induces trivial bundles over every affine cell. In total, this yields a cell decomposition of the quiver Grassmannian $Gr^{\tilde Q}_\tbfe(\tilde P_{m+1}^{(k-1)})$.

@@

\begin{thebibliography}{10}
\bibitem{ass}
Assem, I., Simson, D., Skowronski, A.: Elements of the Representation Theory of Associative Algebras. Cambridge University Press, Cambridge 2007.
\bibitem{bb} Bialynicki-Birula, A.: Some theorems on actions of algebraic groups. Annals of Mathematics \textbf{98}, 480-497 (1973).
\bibitem{bgp}
Bernstein, I.~N., Gelfand, I.~M., Ponomarev, V.~A.: Coxeter functors, and Gabriel's theorem. Russian Mathematical Surveys \textbf{28}(2), 17-32 (1973).
\bibitem{brenner-butler} S. Benner, M.~C.~R. Butler, The equivalence of certain functors occuring in the representation theory of artin algebras and species, J. London Math. Soc., 14 (1976), 183-187.
\bibitem{cc}
Caldero, P., Chapoton, F.: Cluster algebras as {H}all algebras of quiver representations.
Commentarii Mathematici Helvetici \textbf{81}(3), 595-616 (2006).
\bibitem{ck}
  P. Caldero, B. Keller: From triangulated categories to cluster algebras II.  Ann. Sci. \'Ecole Norm. Sup. (4) \textbf{39} (2006), no. 6, pp.~983--1009.
\bibitem{cefr} Cerulli Irelli, Esposito, Franzen, Reineke: Topology of Quiver Grassmannians. Preprint 2017.
\bibitem{cr}
Caldero, P., Reineke, M.: On the quiver Grassmannian in the acyclic case.
Journal of Pure and Applied Algebra \textbf{212}(11), 2369-2380 (2008).
\bibitem{cj} Crawley-Boevey, W., Jensen, Bernt Tore: A note on sub-bundles of vector bundles. Glasgow Mathematical Journal \textbf{48}, 459-462 (2006).

\bibitem{gab} Gabriel, P.: The universal cover of a finite-dimensional algebra. Representations of algebras. Lecture Notes in Mathematics {\bf 903}, 68-105 (1981).
\bibitem{hr} Happel, D., Ringel, C.M.: Tilted Algebras. Transactions of the American Mathematical Society {\bf 274}, no.2, 399-443 (1982).
\bibitem{llz} K. Lee, L. Li, A. Zelevinsky: Greedy elements in rank 2 cluster algebras. Selecta Math. \textbf{20} (2014) pp.~57--82.
\bibitem{rin} Ringel, C.M.: Reflection functors for hereditary algebras. Journal of the London Mathematical Society (2) {\bf 21}, no. 3, 465-479 (1980).
%\bibitem{mats} Matsumura, H.: Commutative ring theory. 
\bibitem{rupel} Rupel, D.: Rank Two Non-Commutative Laurent Phenomenon and Pseudo-Positivity, arXiv.
